---
title: "Cars Consumption Analysis"
author: "Pedro A. Alonso Baigorri"
output:
  html_document: default
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## 1 - Executive Summary

The purpose of this analysis is exploring the relationship between a set of variables and miles per gallon (MPG) (outcome) in the mtcars dataset included in R Studio. On this sense the objective is to answer the following two questions:

* "Is an automatic or manual transmission better for MPG"
* "Quantify the MPG difference between automatic and manual transmissions"

Regression Modelling techniques are used to perform this analysis.

## 2 - Exploratory Data Analysis

With the following r Code I can see the main variables of the dataset.

```{r}
data(mtcars)
head(mtcars, 1)
```
The outcome of this analysis is the variable mtcars$mpg (Miles per Gallon). Higher values
of this variable means lower fuel consumption.

In the annex can be found a plot showing that there is a main dependence of the consumption depending on the transmission type.
Cars with automatic transmission in general consumes more than manual transmission. Other plots with the analysis of the rest of variables can be found also in the annex.

## 3 - Model selection

Now I wil fit a regression model to quantify the impact of Transmission in the consumption but linked also with the rest
of the variables. First of all I'll try a single model using only am as a regressor since it's thekey variable for this analysis

```{r}
fit0 <- lm(mpg ~ factor(am), data = mtcars)
summary(fit0)$coefficients
summary(fit0)$r.squared 
```
As a result we can see that the P-Values are very small, so it means that the relation between mpg and am is very relevant. 
However, the Adjusted R-Squared is very low, meaning that this model only explains the `r round(summary(fit0)$r.squared * 100, 2)` % of the variance of the dataset.I will create two more models adding more variables grouping them according to the correlation index.

```{r}
cor(mtcars)[-1, 1]
```
So for the second model I will use the variables with a corr index > 0.7 (cyl, disp, wt, hp), 
and for the third model the ones between 0.5 - 0.7 (drat, vs, carb). And then I will  use the nested modelling technique to analyze what models to use.

```{r}
fit0 <- lm(mpg ~ factor(am), data = mtcars)
fit1 <- lm(mpg ~ factor(am) + cyl + disp + wt + hp, data = mtcars)
fit2 <- lm(mpg ~ factor(am) + cyl + disp + wt + drat + vs + carb, data = mtcars)
anova(fit0, fit1, fit2)
```
According to this the P-Value for the fit1 model is very low so we keep this model, however the fit2 is not necessary to include it. So the statistics for the final model are:

```{r}
summary(fit1)$coefficients
```
And in this case we have a good Adjusted R-Squared: `r round(summary(fit1)$r.squared * 100, 2)`. The 2nd coef gives the increase in MPG for Manual transmission.

## 4 - Model Residuals

In the annex can be found a figure with the residuals obtained with this model.According to this figure the residuals are very distributed across all the variability of the prediction.

## 5 - Conclusions

After this analysis we can conclude that:

* There is a relationship between Transmission and Consumption: Manual transmission is better for MPG (MPG bigger means fuel consumption higher) 

* We can quantify the difference as an estimation: According to our model cars with manual transmission have a `r fit1$coef[2]` Miles per Gallon more than cars with automatic transmission.

# Annexes

## A1 - Figures of data exploration


```{r echo = FALSE, fig.width=8 , fig.height=4}
par(mfrow=c(1,2))
boxplot(mpg ~ am, data = mtcars, main = "Miles per Gallon Vs Transmission "
        , xlab = "Transmission (0-auto, 1-manual)",  ylab = "MPG")
```

```{r echo = FALSE, fig.width=8 , fig.height=4}
par(mfrow=c(1,2))
boxplot(mpg ~ cyl, data = mtcars, main = "Miles per Gallon Vs Number of Cylinders "
        , xlab = "Cylinders",  ylab = "MPG")
boxplot(mpg ~ vs, data = mtcars, main = "Miles per Gallon Vs Engine type "
        , xlab = "Engine (0-V enginge, 1-S engine)",  ylab = "MPG")
boxplot(mpg ~ gear, data = mtcars, main = "Miles per Gallon Vs Number of Gears "
        , xlab = "Number of Gears",  ylab = "MPG")
boxplot(mpg ~ carb, data = mtcars, main = "Miles per Gallon Vs Number of Carburetors "
        , xlab = "Number of Carburetors",  ylab = "MPG")
```

```{r echo = FALSE, fig.width=8 , fig.height=4}
par(mfrow=c(1,2))
plot(mtcars$disp, mtcars$mpg, main = "Miles per Gallon Vs Gas Displacement "
     , xlab = "Displacement (cu.in.)",  ylab = "MPG")
plot(mtcars$hp, mtcars$mpg, main = "Miles per Gallon Vs Gross Horsepower "
     , xlab = "Horsepower",  ylab = "MPG")
plot(mtcars$drat, mtcars$mpg, main = "Miles per Gallon Vs Rear Axle Ratio "
     , xlab = "Rear Axle Ratio",  ylab = "MPG")
plot(mtcars$wt, mtcars$mpg, main = "Miles per Gallon Vs Weight "
     , xlab = "Weight (1000 lbs)",  ylab = "MPG")
```

## A2 - Residuals plot

```{r echo = FALSE, fig.width=8 , fig.height=4}
residuals <- resid(fit1)
predicted <- predict(fit1)
plot(predicted, residuals, main ="Residuals Vs Predicted")

```



